Valuing Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach

### 2024-11-28 paper/paper.qmd
Adding more charts

```{r}
#| output: asis

data <- read_parquet("../data/cleaned_data/cleaned_data.parquet") |>
  as.data.frame()

@@

```{r #tbl-summ}
modelsummary(list(
  "Final Model" = final_model), output = "kableExtra",
  gof_omit = "BIC|AIC") |>
  column_spec(2, width = "2.5in")
```

@@

testing_data <- testing(split)
```

@@

```{r fig-scatter, fig.cap='Large Uncertainty in the Values of Luxury Homes and Mansions'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
  geom_point(color = "#519C7C", size = 1) +
  geom_function(fun = function(x){x}, linetype = "dashed") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Predicted Price ($ millions)", y = "Actual Price ($ millions)")
```

@@

```{r fig-col, fig.cap='Highest Average Predicted Listing Price in 85 Acre Gated Community'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  select(neighbourhood, predicted_price) |>
  group_by(neighbourhood) |>
  summarise(mean = mean(predicted_price / 10^6)) |>
  arrange(desc(mean)) |>
  head(n = 5) |>
  ggplot(aes(x = reorder(neighbourhood, mean), y = mean)) +
  geom_col(position = "dodge", fill = "#519C7C", color = "black") +
  scale_y_continuous(breaks = seq(0, 5, 0.5)) +
  coord_flip() +
  labs(x = "Neighbourhood", y = "Mean Predicted Price ($ millions)")
```

@@

```{r fig-box, fig.cap='Single Family Homes Valued Significantly Higher Than Other Property Types'}
testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data)) |>
  drop_na() |>
  ggplot(aes(x = as.character(property_type), y = predicted_price / 10^6)) +
  geom_boxplot(fill = "#519C7C", width = 0.25) +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 8, 1)) +
  labs(x = "Property Type", y = "Price ($ millions)")
```

>>>>

```{r, output='asis'}
data <- read_parquet("../data/cleaned_data/cleaned_data.parquet") |>
  as.data.frame()

@@

```{r #tbl-summ}
modelsummary(list(
  "Final Model" = final_model), output = "kableExtra",
  fmt = "%.2f",
  gof_omit = "BIC|AIC") |>
  column_spec(2, width = "2.5in")
```

@@

testing_data <- testing(split)

testing_data <- testing_data |>
  mutate(predicted_price = predict(final_model, newdata = testing_data),
         predicted_price_sqft = predicted_price / sqft) |>
  drop_na()
```

@@

```{r fig-scatter, fig.cap='Large Uncertainty in the Values of Luxury Homes and Mansions'}
testing_data |>
  ggplot(aes(x = predicted_price / 10^6, y = price / 10^6)) +
  geom_point(color = "#519C7C", size = 1) +
  geom_function(fun = function(x){x}, linetype = "dashed") +
  scale_x_continuous(breaks = seq(0, 20, 2)) +
  scale_y_continuous(breaks = seq(0, 20, 2)) +
  labs(x = "Predicted Price ($ millions)", y = "Actual Price ($ millions)")
```

@@

```{r fig-col, fig.cap="Highest Average Predicted Listing Price in Seattle's 85 Acre Gated Community"}
testing_data |>
  select(neighbourhood, predicted_price) |>
  group_by(neighbourhood) |>
  summarise(mean = mean(predicted_price / 10^6)) |>
  arrange(desc(mean)) |>
  head(n = 5) |>
  ggplot(aes(x = reorder(neighbourhood, mean), y = mean)) +
  geom_col(position = "dodge", fill = "#519C7C", color = "black") +
  scale_y_continuous(breaks = seq(0, 5, 0.5)) +
  coord_flip() +
  labs(x = "Neighbourhood", y = "Mean Predicted Price ($ millions)")
```

@@

```{r fig-box, fig.cap='40 Year Old Homes Among the Lowest in Actual and Predicted Values'}
testing_data |>
  group_by(property_age) |>
  summarise(mean_price = mean(price) / 10^6,
            mean_predicted = mean(predicted_price) / 10^6) |>
  pivot_longer(names_to = "price_type", cols = c(mean_price, mean_predicted),
               values_to = "value") |>
  mutate(price_type = recode(price_type, 
                             "mean_price" = "Actual", 
                             "mean_predicted" = "Predicted")) |>
  ggplot(aes(x = property_age, y = value, color = price_type)) +
  geom_point(size = 0.75) +
  scale_color_manual(values = c("black", "#519C7C")) +
  geom_smooth(method = "lm", formula = y ~ I(x^2) + x, se = FALSE) +
  facet_wrap(vars(price_type)) +
  theme(legend.position = "top") +
  labs(x = "Property Age", y = "Average Price ($ millions)", color = "Price Type")
```
